---
title: "R Notebook"
output: html_notebook
---

## Spark Demo

```{r setup}
library(ggplot2)
library(dplyr)
library(sparklyr)
```


First we connect to spark (see the spark tab). In this case we'll run spark locally. 
```{r connect}
sc <- spark_connect(master="local")
spark_web(sc)
```

We'll copy into spark the flights dataset (24 MB)
```{r}
copy_to(sc, nycflights13::flights, "flights")
#spark_web(sc)
```

We use dplyr syntax to filter, mutate, and summarize the data in spark, collect forces the execution and returns the data to R
```{r}
flights <- tbl(sc, "flights")

delay <- flights %>% 
  select(arr_delay,dep_delay, distance) %>% 
  filter(arr_delay > 15) %>% 
  mutate(in_air_delay = arr_delay - dep_delay) %>% 
  group_by(distance) %>% 
  summarize(avg_in_air_delay=mean(in_air_delay)) %>%
  collect()

#spark_web(sc)
```


Now we can use R's strength in visualization ...
```{r}
library(ggplot2)
ggplot(data=delay, aes(x=distance, y=avg_in_air_delay))+geom_point() + geom_smooth(method=loess)
```

Or interactively
```{r}
library(highcharter)
hchart(delay$avg_in_air_delay)
```

We can also fit a number of ML algorithms. For example, we can fit a linear model in Spark, and then inspect the results in R.

```{r}
model <- 
  flights %>% 
  filter(arr_delay >0, distance>0) %>% 
  ml_linear_regression(response="arr_delay", features=c("distance"))

summary(model)
```


